Multi-state modelling with R: the {\tt msm} package \vskip 0.2in \large{Version <<echo=FALSE,results=tex>>= cat(version) @ \vskip 0.1in <<echo=FALSE,results=tex>>= cat(format(Sys.time(), " @ }

### 2013-12-05 pkg/inst/doc/msm-manual.Rnw
Don't evaluate odds.msm call in manual, see if this fixes build
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/msm@31 edb9625f-4e0d-4859-8d74-9fd3b1da38cb
Don't evaluate fev2.msm, see if this fixes the build
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/msm@32 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\item[Odds ratios for misclassification] The function
  \Rfunction{odds.msm} gives the estimated odds ratios corresponding to
  each covariate effect on the misclassification probabilities.

<<>>=
odds.msm(cavmiscsex.msm)
@

@@

<<>>=
hmodel2 <- list(hmmMETNorm(mean=90, sd=16, sderr=8,
                           lower=80, upper=Inf, meanerr=0),
                hmmMETNorm(mean=54, sd=18, sderr=8,
                           lower=0, upper=80, meanerr=0),
                hmmIdent(999))

@@

sojourn.msm(fev2.msm)
@

@@

The following plot illustrates a trajectory of declining FEV$_1$ from
the first lung transplant recipient in this dataset.  This is produced
by the following R code.  The Viterbi algorithm is used to locate the
most likely point at which this individual moved from BOS state 1 to
BOS state 2, according to the fitted Model 2.  This is illustrated by
the vertical dotted line.  This is the point at which the individual's
lung function started to remain consistently below 80\% baseline
FEV$_1$.

<<fig=TRUE>>=
keep <- fev$ptnum==1 & fev$fev<999
plot(fev$days[keep], fev$fev[keep], type="l",
ylab=expression(paste("% baseline ", FEV[1])), xlab="Days after transplant")
vit <- viterbi.msm(fev2.msm)[keep,]
(max1 <- max(vit$time[vit$fitted==1]))
(min2 <- min(vit$time[vit$fitted==2]))
abline(v = mean(max1,min2), lty=2)
text(max1 - 500, 50, "STATE 1")
text(min2 + 500, 50, "STATE 2")
@

>>>>

\item[Odds ratios for misclassification] The function
  \Rfunction{odds.msm} would give the estimated odds ratios corresponding to
  each covariate effect on the misclassification probabilities.

\begin{Scode}
odds.msm(cavmiscsex.msm)
\end{Scode}

@@

\begin{Scode}
hmodel2 <- list(hmmMETNorm(mean=90, sd=16, sderr=8,
                           lower=80, upper=Inf, meanerr=0),
                hmmMETNorm(mean=54, sd=18, sderr=8,
                           lower=0, upper=80, meanerr=0),
                hmmIdent(999))

@@

sojourn.msm(fev2.msm)
\end{Scode}

@@

The following code will create a plot that illustrates a trajectory of declining FEV$_1$ from
the first lung transplant recipient in this dataset.  The Viterbi algorithm is used to locate the
most likely point at which this individual moved from BOS state 1 to
BOS state 2, according to the fitted Model 2.  This is illustrated by
a vertical dotted line.  This is the point at which the individual's
lung function started to remain consistently below 80\% baseline
FEV$_1$.

\begin{Scode}
keep <- fev$ptnum==1 & fev$fev<999
plot(fev$days[keep], fev$fev[keep], type="l",
ylab=expression(paste("% baseline ", FEV[1])), xlab="Days after transplant")
vit <- viterbi.msm(fev2.msm)[keep,]
(max1 <- max(vit$time[vit$fitted==1]))
(min2 <- min(vit$time[vit$fitted==2]))
abline(v = mean(max1,min2), lty=2)
text(max1 - 500, 50, "STATE 1")
text(min2 + 500, 50, "STATE 2")
\end{Scode}
